---
title: "11. Data Reformatting"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Data for this analysis: [Mike Byerly. Alaska commercial salmon catches by management region (1886- 1997). Gulf of Alaska Data Portal. df35b.304.2.](https://knb.ecoinformatics.org/view/df35b.304.2). 

Tutorial for this analysis from Arctic Data Center [here](http://training.arcticdata.io/materials/arctic-data-center-training/data-cleaning-and-manipulation.html). 

```{r, warning=FALSE, message=FALSE}
# Load libraries
library(dplyr)
library(tidyr)

# to call stats filter - stats::filter()

```


# Reformat Catch Data

* Remove "all" column 
* Create species column
    - Move from wide to long 
* General QA 

```{r}

# Can use DataONE package to check and see if there's a new version of the data available! 

catch_original <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method="libcurl"), stringsAsFactors=FALSE)

head(catch_original)

```

Remove the `all` and `notesRegCod` columns using `select`. 

Cmd + shift + m is the shortcut for the pipe operator. 

```{r}

# Key == Species
# Value == Catch 

# Unless you specify otherwise, gather will translate all values in data frame
# into a key value pair

catch_long <- catch_original %>%  
  select(-All, -notesRegCode) %>% # Select columns that we want
  # Gather over everything except region and year
  # Moves from wide -> long 
  gather(key="species", value="catch", -Region, -Year)

head(catch_long)

```

Convert data table back to wide. 

```{r}

catch_wide <- catch_long %>% 
  spread(key=species, value=catch)

head(catch_wide)

```
## Clean up our data 

* Rename catch to catch_thousands
* Change catch column to be numeric 
* Create new catch column in units num. of fish

```{r}

catch_clean <- catch_long %>% 
  rename(catch_thousands = catch) %>% # new name = old name 
  mutate(catch_thousands = ifelse(catch_thousands == "I", 1, catch_thousands)) %>% 
  mutate(catch_thousands = as.numeric(catch_thousands)) %>% 
  mutate(catch = catch_thousands*1000) %>% # creates new col or changes values in existing col
  select(-catch_thousands)

head(catch_clean)

```

```{r, eval=FALSE, echo=FALSE}

# Look for where as.numeric failed. 
length(which(is.na(catch_long$catch)))
i <- which(is.na(catch_long$catch))
catch_long[i,]

```

## Split - Apply - Combine 

* Calculate mean catch by species 
```{r}

species_mean <- catch_clean %>% group_by(species, Region) %>%  
  summarize(catch_mean = mean(catch), num_obs = n())

species_mean_wide <- spread(species_mean, key=species, value=catch_mean)

year_mean_all <- catch_clean %>% group_by(Year) %>%  
  summarize(catch_mean = mean(catch), num_obs = n()) %>% 
  arrange(-catch_mean) %>%  # Orders in descending order 
  filter(Year >= 1990)

year_mean_byspp <- catch_clean %>% group_by(Year, species) %>%  
  summarize(catch_mean = mean(catch), num_obs = n()) %>% 
  arrange(-catch_mean) %>%  # Orders in descending order 
  filter(Year >= 1990)

year_mean_wide <- spread(year_mean_byspp, key=species, value=catch_mean)

```


# Join the Region Definitions 

```{r}

# Read in regions file 
region_defs <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1",
                            method = "libcurl"),
                        stringsAsFactors = FALSE) %>% 
    select(code, mgmtArea)

# head(region_defs)

```
```{r}

# Left join 
catch_joined <- left_join(catch_clean, region_defs, by=c("Region" = "code"))

# Inner join has same number of rows as left join 
# This indicates that no there are no missing 
# catch_joined <- inner_join(catch_clean, region_defs, by=c("Region" = "code"))

```


# Misc. Functions 

Split and unite. 

```{r}

dates_df <- data.frame(date = c("5/24/1930",
                                "5/25/1930",
                                "5/26/1930",
                                "5/27/1930",
                                "5/28/1930"),
                       stringsAsFactors = FALSE)

dates_split <- dates_df %>% 
  separate(date, c("month", "day", "year"), "/")

dates_unite <- dates_split %>% unite(col = date_iso, year,  month, day, sep="-")

# Make two digit month 
dates_2digmonth <- dates_split %>% mutate(month = str_pad(month, 2, side="left", pad="0"))

# Lubridate package is best for dates! 

```

Save data output. 

```{r}

write.csv(catch_clean, "catch_clean.csv", row.names = F)

```

