---
title: "sf Package Demo"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warnings=FALSE, messages=F}
library(sf)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
library(leaflet)
```

Read in the data. 

```{r}
ak_shp_sf <- read_sf("shapefiles/ak_regions_simp.shp")
```

Take a first look at the data and check the coordinate reference system. 

```{r}
plot(ak_shp_sf)

st_crs(ak_shp_sf)
```

Reproject the data to a better projection (Alaska Albers Equal Area). 

```{r}
ak_shp_sf <- ak_shp_sf %>% 
  st_transform(crs = 3338)

st_crs(ak_shp_sf)
```

Now it looks a lot better when plotted. 

```{r}
plot(ak_shp_sf)
```

## Spatial joins

Read in the population csv file. 
Right now the population data frame doesn't have any region information associated with it, but we want it to. 

```{r}
pop <- read.csv("shapefiles/alaska_population.csv")

# Make it spatial and transform it 
pop_sf <- st_as_sf(pop, 
                   coords=c("lng","lat"),
                   crs=4326, # WGS 84 datum lat lon (we're making an assumption this is the datum)
                   remove=FALSE) %>% 
  st_transform(crs=3338)

```

Assign the points to a region using a join. 

```{r}
# A spatial join is dictated by a function based on the geometries of the objects to be joined 
pop_joined_sf <- st_join(pop_sf, ak_shp_sf, join=st_within)

plot(pop_joined_sf["region"])
```

## Group by and summarize 

Need to aggregate the points within a region to calculate population by region. 

```{r}
pop_region <- pop_joined_sf %>% 
  as.data.frame() %>% 
  group_by(region) %>% 
  summarize(total_pop = sum(population))

head(pop_region)
```

Join population totals to reion shapefiles. 

```{r}
ak_pop_sf <- left_join(ak_shp_sf, pop_region)

plot(ak_pop_sf["total_pop"])
```

With the sf package you can group_by and summarize across geographies. 

```{r}
ak_mgmt <- ak_pop_sf %>% 
  group_by(mgmt_area) %>%
  summarize(total_pop = sum(total_pop))

# By default the summarize function will union across the geometries 
# Add do_union=F to no do the union 

plot(ak_mgmt["total_pop"])
```

Write the shapefile. 

```{r}
write_sf(ak_shp_sf, "shapefiles/ak_regions_population.shp")
```

# Static plots 

```{r}
ggplot() + 
  geom_sf(data=ak_pop_sf, aes(fill=total_pop)) + # Must have data= call! 
  geom_sf(data=pop_sf, size=0.5) + 
  theme_bw() + 
  labs(fill="Total Population") + 
  scale_fill_continuous(low="khaki", high="firebrick", labels=comma)

# labels=comma is from the scales library
```

## Interactive maps

Leaflet, unlike ggplot2, wants to project your data for you. 

In order to work with leaflet, you have to give it an unprojected file. 

```{r}

# Create leaflet projection object 
epsg3338 <- leaflet::leafletCRS(
  crsClass = "L.Proj.CRS",
  code = "EPSG:3338",
  proj4def =  "+proj=aea +lat_1=55 +lat_2=65 +lat_0=50 +lon_0=-154 +x_0=0 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs",
  resolutions = 2^(16:7))

```

```{r}
# Unproject our data 
ak_crs <- st_transform(ak_pop_sf, crs=4326) #unprojected data with WGS84 datum
pop_crs <- st_transform(pop_sf, crs=4326) 
```

```{r}

pal <- colorNumeric(palette="Reds", domain=ak_crs$total_pop) # Matches domain values to colors

m <- leaflet(options = leafletOptions(crs = epsg3338)) %>% 
  addPolygons(data = ak_crs, 
              fillColor= ~pal(total_pop),
              weight=1, 
              label= ~region, 
              fillOpacity =1) %>% 
  addCircleMarkers(data=pop_crs, 
                   lat= ~lat,
                   lng= ~lng, 
                   radius= 2, 
                   fillColor="gray",
                   fillOpacity=1,
                   weight=0.25, 
                   color="black",
                   label= ~pop_sf$city)

m
```









